---
title: "About ABM"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    theme:
      base_font: !expr bslib::font_google("Open Sans")
    toc: true
---


```{r setup2, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Background

## What is the aim of ABM at KTH?

The Annual Bibliometric Monitoring at KTH (shortened to ABM) contains statistics on publication output, citation impact and co-publishing. The ABM, together with goals regarding publishing and citation will be a part of the schools’ operational agreements and to be followed up in the school dialogues. It can be used as a basis for understanding how one’s own activities contribute to this goal. 

## Which publications are included in ABM?

Only publications registered in DiVA that researchers have published when working at KTH are included. 

Only KTH-affiliated publications are taken into account.

## At what organizational levels is it available?

The ABM is available at different levels, for KTH in total, schools, departments and individual researchers. Results for the schools and departments are open for all, but to see the individual ABM you will need to login. The report is the same at all levels.

## What data sources is it based on?

ABM is based on data in BIBMET, which in turn collates data from other reference sources, including Web Of Science and Scopus. Statistics regarding citations and co‑publishing are based on the subset of publications in DiVA that are registered in Web of Science (Citation impact, Journal impact and Co‑publishing tabs) or Scopus (Scopus tab).

The bibliometric indicators referred are based on publications registered in DiVA and published 2013 to 2019. Only publications which have been affiliated to KTH are included. This means that publications written by a researcher before she/he was employed at KTH, and that are not affiliated to KTH, will not be included in the statistics.

# Contact information

If you have questions about the ABM at KTH please contact <mailto:biblioteket@kth.se>

This form can also be used to submit a question:

- https://www.kth.se/biblioteket/anvanda-biblioteket/oppettider-kontakt/kontaktformular-1.863120

Read more about other ways to get in touch here:

- https://www.kth.se/biblioteket/anvanda-biblioteket/oppettider-kontakt

# Accessibility

We aim at providing content in an accessible fashion in compliance with [WCAG 2.1](https://www.w3.org/TR/WCAG21/){target="_blank"}, and are aware that there is room for improvement for some of the content.

If you require access to content not provided in a format that is accessible for you, please contact us at [biblioteket@kth.se](mailto:biblioteket@kth.se). Please find more information about accessibility at KTH here: [KTH Web Accessibility Web Page](https://www.kth.se/gemensamt/tillganglighet){target="_blank"}

# Further information

- [Introduction to bibliometrics at KTH](Bibliometrics.html){target="_blank"}
- [Description of data, methods and indicators in KTH Annual Bibliometric Monitoring](Description_data_and_methods_ABM.pdf){target="_blank"}
- [Information about DiVA and the registration process - Handle publications in DiVA](https://www.kth.se/en/biblioteket/publicera-analysera/hantera-publikationer){target="_blank"}
- [Report on Open Access publishing at KTH](OARapport.pdf){target="_blank"}
- [Formal definitions of field normalized citation indicators at KTH](indicators.html){target="_blank"}

For further questions, contact the KTH Library at [biblioteket@kth.se](mailto:biblioteket@kth.se).

## Attribution for third-party data sources

### Redistribution rights for Web Of Science data

Certain data included in the Annual Bibliometric Monitoring report are derived from the © Science Citation Index Expended (SCIE), Social Sciences Citation Index (SSCI), Arts & Humanities Citation Index (AHCI), Conference Proceedings Citation Index - Sciences (CPCI-S) and Conference Proceedings Citation Index - Social Sciences & Humanities (CPCI -SSH) (2021) of Clarivate Analytics (US) LLC. All rights reserved. No part of these materials may be reproduced, stored in a retrieval system or transmitted in any form or by any means, including electronic, mechanical, photographic, magnetic or other means without the expressed permission of the KTH Library.

<!-- include ABM_guide in this overview/about document -->
```{r child = 'ABM_guide.Rmd'}
```
