---
title: "About ABM"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    theme:
      base_font: !expr bslib::font_google("Open Sans")
    toc: true
---


```{r setup2, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Background

## What is the aim of ABM at KTH?

The Annual Bibliometric Monitoring at KTH (shortened to ABM) contains statistics on publication output, citation impact and co-publishing. The ABM, together with goals regarding publishing and citation will be a part of the schools’ operational agreements and to be followed up in the school dialogues. It can be used as a basis for understanding how one’s own activities contribute to this goal. 

## Which publications are included in ABM?

Only publications registered in DiVA that researchers have published when working at KTH are included. 

Only KTH-affiliated publications are taken into account.

## At what organizational levels is it available?

The ABM is available at different levels, for KTH in total, schools, departments and individual researchers. Results for the schools and departments are open for all, but to see the individual ABM you will need to login. The report is the same at all levels.

## What data sources is it based on?

ABM is based on data in BIBMET, which in turn collates data from other reference sources, including Web Of Science and Scopus. Statistics regarding citations and co‑publishing are based on the subset of publications in DiVA that are registered in Web of Science (Citation impact, Journal impact and Co‑publishing tabs) or Scopus (Scopus tab).

The bibliometric indicators referred are based on publications registered in DiVA and published 2013 to 2019. Only publications which have been affiliated to KTH are included. This means that publications written by a researcher before she/he was employed at KTH, and that are not affiliated to KTH, will not be included in the statistics.

# Contact information

If you have further questions about the ABM at KTH please contact [biblioteket$kth.se](mailto:biblioteket@kth.se) or submit a question using [this form](https://www.kth.se/biblioteket/anvanda-biblioteket/oppettider-kontakt/kontaktformular-1.863120){target="_blank"}.

# Accessibility

We aim at providing content in an accessible fashion in compliance with [WCAG 2.1](https://www.w3.org/TR/WCAG21/){target="_blank"}, and are aware that there is room for improvement for some of the content.

If you require access to content not provided in a format that is accessible for you, please contact us at [biblioteket@kth.se](mailto:biblioteket@kth.se). Please find more information about accessibility at KTH at the [KTH Web Accessibility Web Page](https://www.kth.se/gemensamt/tillganglighet){target="_blank"}.

# Further information
- [Bibliometrics (KTH Library)](https://www.kth.se/en/biblioteket/publicera-analysera/bibliometri/bibliometri-1.1159501){target="_blank"}
- [Introduction to bibliometrics at KTH](Bibliometrics.html){target="_blank"}
- [Information about DiVA and the registration process - Handle publications in DiVA](https://www.kth.se/en/biblioteket/publicera-analysera/hantera-publikationer){target="_blank"}
- [Report on Open Access publishing at KTH](OARapport.pdf){target="_blank"}
- [Formal definitions of field normalized citation indicators at KTH](indicators.html){target="_blank"}

For further questions, contact the KTH Library at [biblioteket@kth.se](mailto:biblioteket@kth.se).

## Attribution for third-party data sources

### Redistribution rights for Web Of Science data

Certain data included in the Annual Bibliometric Monitoring report are derived from the © Science Citation Index Expended (SCIE), Social Sciences Citation Index (SSCI), Arts & Humanities Citation Index (AHCI), Conference Proceedings Citation Index - Sciences (CPCI-S) and Conference Proceedings Citation Index - Social Sciences & Humanities (CPCI -SSH) (2021) of Clarivate Analytics (US) LLC. All rights reserved. No part of these materials may be reproduced, stored in a retrieval system or transmitted in any form or by any means, including electronic, mechanical, photographic, magnetic or other means without the expressed permission of the KTH Library.

### Open Access data

Open Access status of publications is retrieved from [Unpaywall](https://unpaywall.org/).

<!-- ### Scopus

(placeholder) --> <!--  - do we have an attribution text from Scopus? --> 

## General denotation and description of bibliometric indicators at KTH

* **P full:**  Full count of publications.
* **P frac:** Fractional count of publications.
* **C full:** Full count of citations.
* **C frac:** Fractionalized count of citations.
* **C3 frac:** Fractionalized count of citations with a three year window. The indicator shows the number of citations received from publications published the same year and the two following years. Since older publications have had more time to get cited than new publications, time series of citations counts using an open window are not appropriate. By using a fixed citation window of three years comparisons over time can be made.
* **WoS coverage:** Share of publications covered by Web of Science
* **cf:** Mean field normalized citation rate. This indicator normalizes for the variation of citation pattern between subject fields. Each publication is compared to a reference group of publications, i.e. publications within the same Web of Science subject category, published the same year and of the same document type (article or review). This is done by dividing the number of citations of each publication with the average number of citations for the reference group of publications, which results in a normalized citation rate. The indicator value expresses the field normalized average citation rate of the unit’s publications. It follows from the definition of cf that the mean field normalized citation rate of all records in the Web of Science database is 1. A citation rate above one indicates
that the set of publications is cited above world average, e.g. a citation rate of 1.2 indicates that the publications are cited 20 percent above world average.
* **Ptopx%:** The number or share of publications among the x percent most cited. The same reference group as for the field normalized citation rate is used for the indicator. Publications can partly belong to the x percent most cited publications if several publications have the same citation value as the percentile limit or if the publication has been classified into multiple fields with different percentile limits.
* **jcf:** Mean field normalized citation rate for journals and serials. This indicator shows the citation impact of the journals in which the unit has published. It is calculated as an average of the field normalized citation rate of the set of journals in which the analyzed unit has published. If the unit has published multiple articles in the same journal, the journal’s citation rate is counted one time for each article. This journal indicator is normalized for field differences by the same principles as the field normalized citation rate (cf). For a publication published in the year x, the value of the journal is based on the years x-1 to x-5.
* **JtopX%:** The number or share of publications that have been published in journals which are among the X percent most cited. The same reference group as for the journals’ field normalized citation rate (jcf) is used for the indicator. The journals in the topX% category publish X percent of the publications in the reference group. A journal can partly belong to the topX% percent if it stretches over the percentile limit or if it has been classified into multiple fields with different percentile limits.
* **International co-publishing:** This indicator shows the number of publications that has been co-published between two or more countries. The default presentation of this indicator is by full counts.
* **Swe. non-univ. co-publishing:** This indicator shows the number of publications that has been co-published between a university and Swedish non-university organizations. The default presentation of this indicator is by full counts.

<!-- include ABM_guide in this overview/about document -->
```{r child = 'ABM_guide.Rmd'}
```
